---
title: "Section 2.2 Practice: Simulation"
author: "Khaled Alhudhain"
output: pdf_document
---

Use simulation (particularly sample() and replicate()) to solve the following problems. **While these problems are solvable theoretically, you must use simulation to solve them here.** Please reach out if you are uncertain how to handle any RStudio particulars to translate your approach into workable code.

# 1. Your department is holding a raffle. They sell 25 tickets and offer five prizes. They place the tickets in a hat and draw one for each prize. The tickets are sampled without replacement, i.e. the selected tickets are not placed back in the hat. What is the probability of winning a prize if you buy one ticket?


```{r}

set.seed(77)
n_simulations <- 10000 # We will simulate 10000 times

results_without_replacement <- replicate(n_simulations, {
  
  drawn_tickets <- sample(1:25, 5, replace = FALSE)
  1 %in% drawn_tickets
})

p_1_a <- mean(results_without_replacement)
p_1_a
```



# 1b. What is the probability of winning a prize with one ticket if the tickets are sampled *with* replacement?

```{r}
# *with* replacement
set.seed(77)
results_with_replacement <- replicate(n_simulations, {

  drawn_tickets <- sample(1:25, 5, replace = TRUE)
  1 %in% drawn_tickets
})

p_1_b <- mean(results_with_replacement)
p_1_b
```


\newpage
# 2. If 100 balls are randomly placed into 25 urns, estimate the probability that at least one of the urns is empty. _Hint: look at the `length()` and `unique()` functions from chapter 1_.



```{r}
set.seed(77)
n_simulations <- 10000

# Using replicate as it Hinted
results <- replicate(n_simulations, {
  urns <- sample(1:25, 100, replace = TRUE)
  length(unique(urns)) < 25
})

p2 <- mean(results)
p2
```



\newpage
# 3. A standard deck of cards has 52 cards, four each of 2,3,4,5,6,7,8,9, 10,J,Q,K,A. In blackjack, a player gets two cards and adds their values. Cards count as their usual numbers, except Aces are 11 (or 1), while K, Q, J are all 10. Use R to simulate dealing two cards, and compute these probabilities. 

_Hint, create the deck of cards using their numeric value, not the letters "J", "Q", "K" etc.--rep() and : are helpful in doing so._

## a. A **blackjack** means getting an Ace and a value ten card. What is the probability of getting a blackjack from your initial draw?



```{r}
set.seed(777)
deck <- c(rep(2:9, each = 4), rep(10, 16), rep(11, 4)) # Creating the deck

n_simulations <- 10000

results <- replicate(n_simulations, {
  hand <- sample(deck, 2, replace = FALSE)
  sum(hand) == 21 && 11 %in% hand
})

p_3_a<- mean(results)
p_3_a
```



## b. What is the probability of getting 18 from your initial draw? Assume that an Ace counts as 11.


```{r}
set.seed(777)
deck <- c(rep(2:9, each = 4), rep(10, 16), rep(11, 4))

n_simulations <- 10000

results <- replicate(n_simulations, {
  hand <- sample(deck, 2, replace = FALSE)
  sum(hand) == 18
})

p_3_b<- mean(results)
p_3_b
```



\newpage
# 4. Deathrolling in World of Warcraft works as follows: Player 1 tosses a 1000 sided die. Say they get $x_{1}$. Then player 2 tosses a die with $x_{1}$ sides on it. Say they get $x_{2}$. Player 1 tosses a die with $x_{2}$ sides on it. This pattern continues until a player rolls a 1. The player who loses is the player who rolls a 1. 

_Hint, you may want to create a separate variable for each person rolling._

## a. What is the probability a 1 will be rolled on the 4th roll in deathroll? 


```{r}
set.seed(77)
n_simulations <- 10000

results <- replicate(n_simulations, {
  x <- 1000  
  for (i in 1:3) {
    x <- sample(1:x, 1)
    if (x == 1) return(FALSE)  
  }
  x <- sample(1:x, 1) 
  x == 1
})

p_4_a <- mean(results)
p_4_a 
```


## b. What is the probability that the second person will lose ON THEIR FIRST ROLL? 

```{r}
set.seed(77)
n_simulations <- 10000

results <- replicate(n_simulations, {
  x1 <- sample(1:1000, 1) # Player 1 starts by rolling a 1000 sided die
  
  x2 <- sample(1:x1, 1) # P2 rolls a x1 sided one
  
  x2 == 1
})

p_4_b <- mean(results)
p_4_b
```



