---
title: "Ch 3.3-3.6 Practice: Functions of Random Variables and Named Discrete Distributions"
author: "Khaled Alhudhain"
output: pdf_document
---



# 1. (20 pts) Let $X$ be a discrete random variable with pmf given by 
P(X=0)=1/4, P(X=1)=1/2, P(X=2)=1/8, P(X=3)=1/8

## a. Find the pmf of $Y=X-1$.


P(Y=-1) = P(X=0) = 1/4 = 0.25


P(Y=0) = P(X=1) = 1/2 = 0.5


P(Y=1) = P(X=2) = 1/8 = 0.125


P(Y=2) = P(X=3) = 1/8 = 0.125


## b. Find the theoretical pmf of $V=(X-1)^{2}$.
Note that there are two ways to get V = 1, so the pmf needs to be simplified. 


P(V=0) = P(X=1) = 1/2 = 0.5




P(V=1) = P(X=0) + P(X=2) = 1/4+ 1/8 = 0.375



P(V=4) = P(X=3) = 1/8 = 0.125



## c. Calculate the theoretical values for $E(V)$ and $Var(V)$. 



$E(V)$ = 0(P(V=0)) + 1(P(V=1)) + 4(P(V=4)) = 0 + 0.375 + 4(0.125) = 0.875



$E(V^2)$ = (0^2)(P(V=0)) + (1^2)(P(V=1)) + (4^2)(P(V=4)) = 0 + 0.375 + 16(0.125) = 2.375



$Var(V)$ = $E(V^2)$ - $(E(V))^2$ = 2.375 - (0.875)^2 = 103/64 = 1.609375




## d. Estimate $E(V)$ and $Var(V)$ using simulation. 

```{r}

set.seed(777)

n <- 100000

X <- sample(0:3, n, replace = TRUE, prob = c(1/4, 1/2, 1/8, 1/8))

V <- (X - 1)^2

E_V <- mean(V)

Var_V <- var(V)

cat("Simulated E(V) = ", E_V, "\n")
cat("Simulated Var(V) = ", Var_V, "\n")
```




\newpage
# 2. (10 pts) Let $X$ be a random variable with mean $\mu$ and standard deviation $\sigma$. Use the properties of the variance to find the mean and standard deviation of $\frac{X-\mu}{\sigma}$.

Mean:  E[$\frac{X-\mu}{\sigma}$] = ${\frac{1}\sigma}$E[{$X-\mu$}] = ${\frac{1}\sigma}$(E[X] - $\mu$ ) = ${\frac{1}\sigma}$($\mu$ - $\mu$ ) = 0


Variance: Var[$\frac{X-\mu}{\sigma}$] = ${(\frac{1}\sigma)^2}$Var[{$X-\mu$}] = ${(\frac{1}\sigma)^2}$(Var[X] - Var[$\mu$] ) = ${(\frac{1}\sigma)^2}$(Var[X] - 0 ) = ${(\frac{1}\sigma)^2}$($\sigma^2$) = 1

\newpage
# 3. (20 pts) A recent national study showed that approximately 45\% of college students binge drink. Let $X$ equal the number of students in a random sample of size $n=12$ who binge drink. Using both the pmf by hand, and R commands, find the probability that

## a. $X$ is at most 2.



P(X<=2) = P(X = 0) + P(X = 1) + P(X = 2) = (0.55^12)C(12,0) + (0.55^11) (0.45^1) C(12,1) + (0.55^10) (0.45^2) C(12,2) = 0.0421419826



## b. $X$ is at least 1.


P(X >= 1) = 1 - P(X = 0) = 1 - (0.55^12)C(12,0) = 0.999233782


## c. Use simulation to obtain the mean and variance of $X$. 
```{r}
set.seed(777)

n_sim <- 100000

n <- 12
p <- 0.45

X_sim <- rbinom(n_sim, n, p)

mean_X <- mean(X_sim)
var_X <- var(X_sim)

cat("Average of X = ", mean_X, "\n")
cat("Var of X = ", var_X, "\n")
```


\newpage
# 4. (20 pts) There is a .85 probabilty that Avery will show up to work exactly on time (not 1 minute late), independent of any other day. This week, it's crunch time on a huge project, and employees are expected to come in every day of the week. The company gives a $100 bonus for every employee that shows up to work exactly on time at least 5 days in a week. 

## a. Assuming the work week starts on Sunday (as day 1 of 7), what is the probability that Avery gets this bonus by Thursday?


He has to show up every day from Sunday to Thursday since there are 5 days between them inclusive:

P(X = 5) = (0.85)^5 = 0.4437





## b. What day on average will Avery get this bonus?


Binomial distribution: E[X] = n/p = 5/0.85 = 5.88

So, on average, he will get the bonus in about 6 days or on "Sunday + 6 days" = Saturday

## c. What is the probability that Avery won't get this bonus?



P(X < 5) = 1- (P(X = 5) + P(X = 6) + P(X = 7)) = 0.07376516



\newpage

# 5. (10 pts) Flaws in a certain type of drapery material appear on the average of two in 150 square feet.  If we assume a Poisson distribution, find the probability of at most 2 flaws in 450 square feet.



lambda for 150 square feet = 2


So, lamda for 450 square feet = 2*3 = 6



P( X <= 2) = P(X = 0) + P(X = 1) + P(X = 2) = e^(-6) (6^0 / 0!) + e^(-6) (6^1 / 1!) + e^(-6) (6^2 / 2!) = e^(-6) * [ 1 + 6 + 18] = 25(e^-6) = 0.0619688 




\newpage

# 6. (20 pts) In the game of Scrabble, players make words using letter tiles. The tiles consist of 42 vowels and 58 non-vowels.

## a. If a player draws 7 tiles, what is the probability of getting 7 vowels?

Total # of tiles = 42 + 58 = 100


P(7 vowels) = C(42,7) / C(100,7) = 59163 / 35104300 = 0.001685349



## b. If a player draws 7 tiles, what is the probability of 2 or fewer vowels?

X: # of drawn vowels


P(X <= 2) = P(X = 0) + P(X = 1) + P(X = 2) = C(58,7)/C(100,7) + (C(58,6) C(42,1))/C(100,7) + (C(58,5) C(42,2))/C(100,7) = 47415 / 127652 = 0.3714395387


## c. What is the expected number of vowels drawn when drawing 7 tiles?


Hyper-Geometric Dist:


E[X] = n*K/N = 7(42)/100 = 2.94



## d. What is the standard deviation of the number of vowels drawn when drawing 7 tiles?


Hyper-Geometric Dist:

Std = sqrt(n(K/N)((N-K)/N)((N-n)/(N-1))) = sqrt(7(0.42)(0.58)(93 / 99)) = 1.26564



