---
title: "Ch 4.1 Practice: PDFs and CDFs"
author: "Khaled Alhudhain"
output: pdf_document
---

# 1. (30 pts) Let $X$ be a random variable with pdf given by $f(X)=2x$ for $0\leq x\leq 1$ and 0 otherwise.

## a. Find $P(X\geq 1/2)$


$P(X\geq 1/2)$ = integral(2x) from x= 1/2 to x=1 = $x^2$ from x= 1/2 to x=1 = 1^2 - (1/2)^2 = 1 - 1/4 = 3/4 = 0.75



Code:

```{r}
p_1a<- function() {
  integrate(function(x) 2 * x, lower = 0.5, upper = 1)$value
}
p_1a()
```



## b. Find $P(X\geq1/2|X\geq 1/4)$.


$P(X\geq1/2|X\geq 1/4)$ = $P(X\geq1/2 \bigcap X\geq 1/4)  \div P( X\geq 1/4)$ 



$P(X\geq1/2|X\geq 1/4)$ = (x^2 from x= 1/2 to x=1) / (x^2 from x= 1/4 to x=1) = (1^2 - (1/2)^2) / (1^2 - (1/4)^2) = (3/4)/(15/16) = 4/5 = 0.8


code:


```{r}
p_1b <- function() {
  integrate(function(x) 2 * x, lower = 0.25, upper = 1)$value
}
output_1b <- function() {
  p_1a() /p_1b()
}
output_1b()
```





\newpage
# 2. (20 pts) Let $X$ be a random variable with pdf given below. Find the constant c such that this is a valid pdf. 

$$
f(x)= \Bigg\{
\begin{array}{cc}
cx^{2} & \mbox{for }0 \leq x < 1 \\
c(2-x)^{2} & \mbox{for }1 \leq x \leq 2 \\
0  & otherwise.
\end{array}
$$
Integral of f(x) over all its domain has to be equal to 1.


[inetgral(cx^2) from x = 0 to x = 1] + [inetgral(c(2-x)^2) from x = 1 to x = 2] = c[(x^3)/3  from x = 0 to x = 1] + c[((x-2)^3)/3  from x = 1 to x = 2] = 1

So:

c[1/3] + c[0+1/3] = 1

which means that:

c = 3/2 = 1.5


code:



```{r}
Q_2 <- function() {
  integral1 <- integrate(function(x) x^2, lower = 0, upper = 1)$value
  integral2 <- integrate(function(x) (2 - x)^2, lower = 1, upper = 2)$value
  c <- 1 / (integral1 + integral2)
  return(c)
}
Q_2()
```

\newpage
# 3. (30 pts) Suppose that the p.d.f. of a random variable $X$ is as follows:

$$
f(x)= \Big\{ \begin{array}{cc} 
\frac{1}{8}x & for\ 0\leq x \leq 4 \\
0 & otherwise.
\end{array}
$$

## a. Find the value of $t$ such that P$(X\leq t)=1/4$.


P$(X\leq t)$ = $\int_{0}^{t} (x/8)   dx$ = (x^2)/16 from x = 0 to x = t = $\frac{t^2}{16}$ = 1/4 

which means:


$t^2$  = 4

and since t is non-negative:

t = 2



code:

```{r}
Q_3a <- function() {
  solve <- function(t) (1/16) * t^2 - 1/4
  round(uniroot(solve, lower = 0, upper = 4)$root)
}
Q_3a()

```


## b. Find the value of $t$ such that P($X\geq t)=1/3$.


P($X\geq t$ = $\int_{t}^{4} (x/8)   dx$ = (x^2)/16 from x = t to x = 4 = $\frac{16-t^2}{16}$ = 1/3



which means:


16 - $t^2$  = $\frac{16}{3}$


which gives:


$t^2$ = $\frac{32}{3}$

and since t is non-negative:

t = $\sqrt{\frac{32}{3}}$



code:

```{r}
Q_3b <- function() {
  solve <- function(t) 1 - (1/16) * t^2 - 1/3
  uniroot(solve, lower = 0, upper = 4)$root
}
Q_3b()

```


\newpage
# 4. (20 pts) Let $f(y)=\frac{3}{2}y^{2}$ for $-1\leq y\leq 1$. Find $P(|Y-\frac{1}{2}|<\frac{1}{4})$.


$P(|Y-\frac{1}{2}|<\frac{1}{4})$ = $P(\frac{-1}{4} < Y-\frac{1}{2}<\frac{1}{4})$ = $P(\frac{1}{4} < Y<\frac{3}{4})$


P($\int_{1/4}^{3/4}\frac{3}{2}y^{2} dy$) = $\frac{y^3}{2}$  from x = 3/4 to x = 1/4 = (1/2)[ (27/64) - (1/64) ] = 13/64


code:


```{r}
Q_4 <- function() {
  integrate(function(y) (3/2) * y^2, lower = 1/4, upper = 3/4)$value
}
Q_4()
```

